---
title: "edin_analysis"
output: word_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r setup}
library(tidyverse)
library(car)
library(ggplot2)
library(magrittr)
library(corrplot)
library(rgdal)
library(broom)
library(maptools)
library(wesanderson)
library(factoextra)
library(gridExtra)
library(spgwr)
```

## Load data & Clean data
```{r load_data}
shape_file <- readOGR(dsn = "simd2020_edinburgh/", 
                      layer = "simd2020_edinburgh")

spdf_fortified <- tidy(shape_file) 
spdf_fortified <- spdf_fortified %>% mutate(id = as.integer(id))
dat <- shape_file@data %>% select(Data_Zone, DEPRESS, ALCOHOL, 
                                  crime_rate, employment,
                             X, Y)
dat <- dat %>% 
  mutate(id = seq(0,nrow(dat)-1)) %>% 
  tibble()


# Data cleaning
dat <- dat %>% 
  tibble() %>% 
  mutate(DEPRESS = as.numeric(sub("%", "", DEPRESS)) / 100,
         employment = as.numeric(sub("%", "", employment)) / 100,
         crime_rate = as.integer(crime_rate),
         ALCOHOL = as.integer(ALCOHOL))

## Remove a value of 0
dat <- dat %>%
  mutate(ALCOHOL = ifelse(ALCOHOL == 0, NA, ALCOHOL),
         crime_rate = ifelse(crime_rate == 0, NA, crime_rate),
         employment = ifelse(ALCOHOL == 0, NA, employment))

```

## Visualize data through boxplots
```{r boxplot}

boxplot_depress <- ggplot(dat, aes(y=DEPRESS)) +
  geom_boxplot() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x=element_blank())

boxplot_crime_rate <- ggplot(dat, aes(y=crime_rate)) +
  geom_boxplot() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x=element_blank())

boxplot_employment <- ggplot(dat, aes(y=employment)) +
  geom_boxplot() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x=element_blank())

boxplot_alcohol <- ggplot(dat, aes(y=ALCOHOL)) +
  geom_boxplot() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x=element_blank())

grid.arrange(boxplot_depress,
             boxplot_alcohol,
             boxplot_crime_rate,
             boxplot_employment,
             nrow = 2)


```

## Remove outliers based on IQR methdology
```{r remove_outlier}

depress_outlier <- boxplot.stats(dat$DEPRESS)$out
crime_outlier <- boxplot.stats(dat$crime_rate)$out
emp_outlier <- boxplot.stats(dat$employment)$out
alcohol_outlier <- boxplot.stats(dat$ALCOHOL)$out

# Replace outlier with NA
dat <- dat %>%
  mutate(DEPRESS = ifelse(DEPRESS %in% depress_outlier, NA, DEPRESS),
         ALCOHOL = ifelse(ALCOHOL %in% alcohol_outlier, NA, ALCOHOL),
         crime_rate = ifelse(crime_rate %in% crime_outlier, NA, crime_rate),
         employment = ifelse(employment %in% emp_outlier, NA, employment))

## create spatial data
spatial_dat <- dat %>% inner_join(spdf_fortified, by = "id")

## Remove NA(outlier)
dat <- dat %>% filter(!is.na(DEPRESS),
                      !is.na(crime_rate), 
                      !is.na(employment), 
                      !is.na(ALCOHOL))
```

## VIF

```{r vif}
lm_mod <- lm(DEPRESS~ crime_rate + employment + ALCOHOL,  data = dat)

# all less than 5
vif(lm_mod)

```

## Data visulization - Correlation map
``` {r correlation}

correlation = cor(dat %>% select(DEPRESS,
                                 ALCOHOL,
                                 crime_rate,
                                 employment))
corrplot(correlation, method = 'color')
```

## Data visulization - Scatter grid
``` {r scatter_grid}
## Scatter grid
pairs(dat %>% select(DEPRESS,
                     ALCOHOL,
                     crime_rate,

                       employment)
        , pch = 19)

```

## Data visulization - Heat Map
``` {r heat_map}
## Heat maps

pal <- wes_palette("Zissou1", 100, type = "continuous")

depress_map <- ggplot() + 
  geom_polygon(data = spatial_dat, aes(x = long,
                                       y = lat, 
                                       group = group,
                                       fill = DEPRESS)) + 
  scale_fill_gradientn(colours = pal)

crime_rate_map <- ggplot() + 
  geom_polygon(data = spatial_dat, aes(x = long,
                                       y = lat, 
                                       group = group,
                                       fill = crime_rate)) +
  scale_fill_gradientn(colours = pal)


alcohol_map <- ggplot() + 
  geom_polygon(data = spatial_dat, aes(x = long,
                                       y = lat, 
                                       group = group,
                                       fill = ALCOHOL)) +
  scale_fill_gradientn(colours = pal)

employment_map <- ggplot() + 
  geom_polygon(data = spatial_dat, aes(x = long,
                                       y = lat, 
                                       group = group,
                                       fill = employment)) + 
  scale_fill_gradientn(colours = pal)

grid.arrange(depress_map,
             alcohol_map,
             crime_rate_map,
             employment_map,
             nrow = 2)
```

## K-means with elbow method
```{r k_means_1}

# scale data
scale_dat <- scale(dat %>% 
                     select(DEPRESS,
                            ALCOHOL,
                            crime_rate,
                            employment))

## Elbow method
fviz_nbclust(scale_dat, kmeans, method = "wss") + 
geom_vline(xintercept = 3, linetype = 2)
```

## K-means cluster map
```{r k_means_2}
## Add cluster info
set.seed(1)
kmean_model <- kmeans(scale_dat, centers = 3, nstart = 25)
dat <- dat %>% 
  mutate(cluster = factor(kmean_model$cluster))

spatial_dat <- spatial_dat %>%
  left_join(dat %>% select(id, cluster))

ggplot() + 
  geom_polygon(data = spatial_dat, aes(x = long,
                                       y = lat, 
                                       group = group,
                                       fill = cluster)) 


```

## K-means cluster profile
- Cluster 1 has the least depression rate and the lowest crime rate, alcohol hospitalization and employment deprivation.
- Cluster 2 has the most depression rate and the highest crime rate, alcohol hospitalization and employment deprivation.
- Cluster 3 is in the middle between Cluster 1 and 2.
- Most areas belong to Cluster 1, which has the least depression rate.
```{r k_means_cluster profile}

## cluster profiles
dat %>% 
  group_by(cluster) %>% 
  summarize(count = n())

depress_plot <- ggplot(data = dat, aes(x = DEPRESS, fill = cluster)) +
  geom_density(alpha = 0.4)
alcohol_plot <- ggplot(data = dat, aes(x = ALCOHOL, fill = cluster)) + 
  geom_density(alpha = 0.4)
crime_plot <- ggplot(data = dat, aes(x = crime_rate, fill = cluster)) + 
  geom_density(alpha = 0.4)
employment_plot <- ggplot(data = dat, 
                          aes(x = employment, fill = cluster)) + 
  geom_density(alpha = 0.4)

grid.arrange(depress_plot, 
             alcohol_plot,
             crime_plot,
             employment_plot,
             nrow = 2)
```

## OLS regression
```{r ols}

summary(lm_mod)

```

## GWR
```{r gwr}

fbw <- gwr.sel(formula = DEPRESS ~ ALCOHOL + crime_rate + employment, 
               data = dat, 
               coords=cbind(dat$X, dat$Y),
               longlat = TRUE,
               adapt=FALSE, 
               gweight = gwr.Gauss, 
               verbose = FALSE)

gwr_model <- gwr(formula = DEPRESS ~ ALCOHOL + crime_rate + employment,
                 coords = cbind(dat$X, dat$Y),
                 data = dat,
                 bandwidth = fbw,
                 hatmatrix = T,
                 se.fit = T)

gwr_model
```